### 2015-11-02 manuscript.tex
Hide figs for now. urls in text for real data

%\documentclass[]{article}
%\usepackage{natbib}
\documentclass[extra,mreferee]{gji}
%\documentclass[extra]{gji}
\usepackage{timet}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{url}
\usepackage[utf8]{inputenc}

\usepackage{todonotes}

% Document metadata
% =================
\newcommand{\Title}{
    Fast non-linear gravity inversion in spherical coordinates:
    application to the South American Moho
}
\newcommand{\Keywords}{
        Moho;
        gravity inversion;
        spherical coordinates;
        tesseroid;
        South America
}
\title[]{\Title}
\author[]{
    Leonardo Uieda$^{1,2}$,
    Valéria C. F. Barbosa$^{2}$
    \\
    $^1$Universidade do Estado do Rio de Janeiro, Rio de Janeiro, Brazil.
    e-mail: leouieda@gmail.com
    \\
    $^2$Observatório Nacional, Rio de Janeiro, Brazil.
}

\usepackage[pdftex,colorlinks=true]{hyperref}
\hypersetup{
    pdftitle={\Title},
    pdfauthor={Leonardo Uieda (leouieda@gmail.com)},
    pdfsubject={},
    pdfkeywords={\Keywords},
    pdfcreator={pdfTeX},
    allcolors=blue,
}

% Useful commands
% ===============

\begin{document}

%\label{firstpage}
\maketitle

\begin{abstract}
\end{abstract}

\noindent\textbf{Key words:} \Keywords

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methodology}

In potential field methods,
we must isolate the target anomalous density distribution prior to modeling and
inversion.
In our case, the target is the relief of the real Moho undulating around a
reference Moho.
We do this by removing all other effects from the gravity observations.
The first correction is to remove the
scalar gravity of an ellipsoidal reference Earth (the Normal Earth),
hereafter denoted as $\gamma$.
This effect is calculated on the same point $P$ where the observation was made
using the closed-form solution presented by \citet{li2001}.
The difference between the observed gravity ($g$) and Normal gravity
is known as the gravity disturbance,
\todo{Seminar figure here}

\begin{equation}
    \delta(P) = g(P) - \gamma(P).
\end{equation}

The disturbance contains only the gravitational effects of density
distributions that are anomalous with respect to the Normal Earth.
This includes all masses above the surface of the ellipsoid (the topography),
the mass deficiency of the oceans,
the mass deficiency of sedimentary basins,
crustal sources (e.g., igneous intrusions, lateral density changes, etc),
heterogeneities below the upper mantle,
and the effect of the difference between the real Moho
topography and the Moho of the Normal Earth.

All effects must either be removed or assumed negligible
to invert for the anomalous Moho.
The gravitational attraction of the topography, oceans, and basins are
calculated in a spherical Earth approximation by forward modeling using
tesseroids
(Fig.~\ref{fig:tesseroid}).
The tesseroid effects are calculated numerically using
Gauss-Legendre Quadrature (GLQ) integration \citep{asgharzadeh2007}.
The accuracy of the GLQ integration is improved by the adaptive discretization
scheme of \citet{uieda2016}.

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/paper/tesseroid-coord-sys}
    \caption{Sketch of a tesseroid (spherical prism) in a geocentric coordinate
        system (X, Y, Z).
        Observations are made at point P with respect to it's local
        North-oriented coordinate system (x, y, z).
        After \citet{uieda2015},
        available at
        \url{http://dx.doi.org/10.6084/m9.figshare.1495525}.
    }
    \label{fig:tesseroid}
\end{figure}

\subsection{Parametrization}

We parameterize the forward problem by discretizing the anomalous Moho
into a grid of $M_{lon} \times M_{lat} = M$ juxtaposed tesseroids.
The true (real Earth) Moho varies in depth
with respect to the Moho of the Normal Earth.
Hereafter we will refer to the depth of the Normal Earth Moho as $z_{ref}$.
In cases where the true Moho is above $z_{ref}$,
the top of the $k$th tesseroid is the Moho depth $z_{k}$,
the bottom is $z_{ref}$, and the density-contrast ($\Delta\rho$) is positive.
If the Moho is below $z_{ref}$, the top of the tesseroid is $z_{ref}$,
the bottom is $z_k$, and $\Delta\rho$ is negative.
\todo{Seminar figure here}

Considering that the density-contrasts of the tesseroids are fixed parameters,
the predicted gravity anomaly of the Moho is a non-linear function of the
parameters $z_k$, $k=1, \ldots, M$,

\begin{equation}
    d_i = f_i(\mathbf{p}),
\end{equation}

\noindent in which $d_i$ is the $i$th element of the $N$-dimensional predicted
data vector $\mathbf{d}$, $\mathbf{p}$ is the $M$-dimensional parameter vector
containing the $M$ Moho depths ($z_k$),
and $f_i$ is the $i$th non-linear function that maps the parameters onto the
data.
The functions $f_i$ are the radial component of the gravitational attraction
of the tesseroid Moho model.

\subsection{Inverse problem}

We wish to estimate the parameter vector $\mathbf{p}$ from a set of observed
gravity anomaly data $\mathbf{d}^o$.
The least-squares estimate is the one that minimizes the data-misfit function

\begin{equation}
    \phi(\mathbf{p}) =
    [\mathbf{d}^o - \mathbf{d}(\mathbf{p})]^T[\mathbf{d}^o - \mathbf{d}(\mathbf{p})].
    \label{eq:data-misfit}
\end{equation}

Function $\phi(\mathbf{p})$ is non-linear with respect to $\mathbf{p}$.
Thus, we can determine its minimum using gradient-based
iterative optimization
methods like Gauss-Newton or Steepest Descent.
Such methods start from an initial estimate $\mathbf{p}^0$ and repeatedly
update the estimate until a minimum is reached.

For the Gauss-Newton method,
the update at the $k$th iteration,
$\mathbf{\Delta p} = \mathbf{p}^{k+1} - \mathbf{p}^k$,
is the solution of the linear system

\begin{equation}
    \mathbf{H}^k\mathbf{\Delta p} = -\mathbf{\nabla\phi}^k,
    \label{eq:gaussnewton}
\end{equation}

\noindent in which
$\mathbf{\nabla\phi}^k$ and $\mathbf{H}^k$ are, respectively,
the gradient vector and the Hessian matrix of $\phi(\mathbf{p})$.

The Steepest Descent method uses only the gradient direction
to update the initial estimate \citep{kelley1987}.
The update at the $k$th iteration is achieved by equating the Hessian
in Eq.~\ref{eq:gaussnewton} to the identity matrix,

\begin{equation}
    \mathbf{\Delta p} = -\mathbf{\nabla\phi}^k.
    \label{eq:steepest}
\end{equation}

\noindent
Thus, it does not require the computation and storage of the Hessian matrix
nor the solution of linear systems.
However, the Steepest Descent method has poor convergence when the
current solution is close to the minimum of the goal function
\citep{kelley1987}.

The gradient vector and the Gauss-Newton approximation of the Hessian matrix
of $\phi(\mathbf{p})$ are, respectively,

\begin{equation}
    \mathbf{\nabla\phi}^k = -2\mathbf{A}^T[\mathbf{d}^o - \mathbf{d}(\mathbf{p}^k)],
    \label{eq:gradient}
\end{equation}

\noindent
and

\begin{equation}
    \mathbf{H}^k \approx 2\mathbf{A}^T\mathbf{A},
    \label{eq:hessian}
\end{equation}

\noindent in which
$\mathbf{A}$ is the Jacobian or sensitivity matrix,

\begin{equation}
    A_{ij} = \dfrac{\partial f_i}{\partial p_j}(\mathbf{p^k}).
    \label{eq:jacobian}
\end{equation}

\subsection{Regularization}

Non-linear inversions for the relief of an interface (like the Moho)
are ill-posed and require additional constraints in the form of
regularization \citep{silva2001}.
A common approach is to use the first-order Tikhonov regularization to impose
smoothness on the solution.
The cost function for smoothness regularization is given by

\begin{equation}
    \theta(\mathbf{p}) = \mathbf{p}^T\mathbf{R}^T\mathbf{R}\mathbf{p},
\end{equation}

\noindent where $\mathbf{R}$ is an $L \times M$ finite-difference matrix
representing the $L$ first-order differences between adjacent tesseroids.

The solution $\hat{\mathbf{p}}$ to the regularized inverse problem is the one that
minimizes the goal function

\begin{equation}
    \Gamma(\mathbf{p}) = \phi(\mathbf{p}) + \mu\theta(\mathbf{p}),
\end{equation}

\noindent
in which $\mu$ is the regularization parameter that controls the balance
between fitting the observed data and obeying the smoothness constraint.

The goal function $\Gamma(\mathbf{p})$ is also non-linear with respect to
$\mathbf{p}$ and can be minimized using the Gauss-Newton or Steepest Descent
methods.
The gradient vector and Hessian matrix of the goal function are, respectively,

\begin{equation}
    \mathbf{\nabla\Gamma}^k =
        -2\mathbf{A}^T[\mathbf{d}^o - \mathbf{d}(\mathbf{p}^k)] +
        2\mu\mathbf{R}^T\mathbf{R}\mathbf{p}^k,
    \label{eq:gradient-regul}
\end{equation}

\noindent and

\begin{equation}
    \mathbf{H}^k = 2\mathbf{A}^T\mathbf{A} + 2\mu\mathbf{R}^T\mathbf{R}.
    \label{eq:hessian-regul}
\end{equation}

\noindent The parameter updates for the regularized Gauss-Newton and Steepest
Descent methods, respectively, then become

\begin{equation}
    \left[\mathbf{A}^T\mathbf{A} + \mu\mathbf{R}^T\mathbf{R}\right]
    \mathbf{\Delta p} =
        \mathbf{A}^T[\mathbf{d}^o - \mathbf{d}(\mathbf{p}^k)] -
        \mu\mathbf{R}^T\mathbf{R}\mathbf{p}^k,
    \label{eq:gaussnewton-regul}
\end{equation}

\noindent and

\begin{equation}
    \mathbf{\Delta p} =
        \mathbf{A}^T[\mathbf{d}^o - \mathbf{d}(\mathbf{p}^k)] -
        \mu\mathbf{R}^T\mathbf{R}\mathbf{p}^k,
    \label{eq:steepest-regul}
\end{equation}

Producing the regularized solution using the above equations is computationally
costly because of two main factors:
(1) the evaluation and storage of the dense $N \times M$ Jacobian matrix
$\mathbf{A}$
and (2) the solution of the resulting $M \times M$ equation system
(not required for Steepest Descent).
In practice, the derivatives in the Jacobian (Eq.~\ref{eq:jacobian})
are often calculated through a first-order finite-difference approximation.
Thus, evaluating $\mathbf{A}$ requires $2\times M \times N$ forward modeling
operations for each iteration of the gradient descent algorithm.
These computations are performed for each iteration of the optimization.

\subsection{Bott's method}

\citet{bott1960} developed an efficient method to determined the basement
relief of a sedimentary basin from gravity observations.
The method requires data on a regular grid of $N_x \times N_y = N$
observations.
The basement relief is then discretized into an equal grid of $M_x \times
M_y = M$ elements with $M_x = N_x$ and $M_y = N_y$.
Bott's iterative method starts with an initial estimate of the basement relief
$\mathbf{p}^0$ equal to the null vector and updates the estimate using the formula

\begin{equation}
    \mathbf{\Delta p} = \dfrac{\mathbf{d}^o - \mathbf{d}(\mathbf{p}^k)}{2\pi G \Delta \rho},
    \label{eq:bott}
\end{equation}

\noindent
in which $G$ is the gravitational constant and $\Delta \rho$ is the basin
density contrast.
The iterative process stops when the inversion residuals
$\mathbf{r} = \mathbf{d}^o - \mathbf{d}(\mathbf{p}^k)$ fall below the assumed noise level
of the data.

\citet{silva2014} showed that Bott's method can be formulated as
a special case of the Gauss-Newton method (Eq.~\ref{eq:gaussnewton})
by setting the Jacobian matrix (Eq.~\ref{eq:jacobian}) to

\begin{equation}
    \mathbf{A} = 2\pi G \Delta \rho \mathbf{I},
    \label{eq:bott-gaussnewton}
\end{equation}

\noindent
in which $\mathbf{I}$ is the identity matrix.
In this framework,
Bott's method uses a Bouguer plate approximation of the gravitational effect of
the relief, $d_i = 2\pi G \Delta\rho z_i$.
The derivative of $d_i$ with respect to the parameter $z_i$ is
$2\pi G \Delta \rho$, thus linearizing the Jacobian matrix.
However, the non-linearity of the predicted data $\mathbf{d}(\mathbf{p}^k)$ is
preserved.

We propose that Bott's method can also be formulated as a special case of the
Steepest Descent method (Eq.~\ref{eq:steepest}) by setting the Jacobian matrix to

\begin{equation}
    \mathbf{A} = \dfrac{1}{4 \pi G \Delta \rho}\mathbf{I}.
    \label{eq:bott-steepest}
\end{equation}

\noindent
In practice, both formulations lead to Eq.~\ref{eq:bott}.
One of the advantages of Bott's method over the traditional Gauss-Newton or
Steepest Descent is eliminating the computation and storage of the dense
Jacobian matrix $\mathbf{A}$.
Furthermore, Bott's method also does not require the solution of equation
systems.
However, a disadvantage of Bott's method is that it suffers from instability
\citep{silva2014}.
A common approach to counter this issue is to apply a smoothing filter after
the inversion to the unstable estimate, as in \citet{silva2014}.

\subsection{Regularized Bott's method in spherical coordinates}

We propose a regularized version of Bott's method to invert for the relief of
the anomalous Moho in spherical coordinates.
Our formulation maintains the regularized solutions
for Gauss-Newton (Eq.~\ref{eq:gaussnewton-regul}) and
Steepest Descent (Eq.~\ref{eq:steepest-regul})
but replaces the full Jacobian matrix with the Bouguer plate approximations
(respectively, Eq.~\ref{eq:bott-gaussnewton} and \ref{eq:bott-steepest}).
This linearizes the Jacobian matrix and reduces it to a sparse diagonal matrix,
thus eliminating the cost of computing and storing $\mathbf{A}$.
Matrix arithmetic operations can be performed efficiently by taking advantage
of the sparse nature of matrices $\mathbf{A}$ and $\mathbf{R}$.
The same is true for solving the equation system in the Gauss-Newton method
(Eq.~\ref{eq:gaussnewton-regul}).
However, the computational cost of forward modeling is still present.
Particularly, forward modeling using tesseroids is more computationally
intensive than using right-rectangular prisms
because of the numerical integration and adaptive discretization.
Benchmarks suggest that
forward modeling accounts for approximately $99\%$
\todo{Check this number}
of the computation time for a Gauss-Newton inversion
(see supplementary material).
\todo{Better reference}
Hence, this formulation allows us to retain the efficiency of Bott's method
while stabilizing the solution through the well established formalism of
Tikhonov regularization.

\subsection{Estimating the regularization parameter}

The regularization parameter $\mu$ controls how much smoothness is applied to
the inversion result.
An optimal value of $\mu$ will stabilize and smooth the solution while not
compromising the fit to the observed data.
Two widely used methods to estimate an optimal $\mu$ are
the L-curve criterion and cross-validation \citep{hansen1992}.
Here, we will adopt the hold-out method of cross-validation \citep{kim2009}.
\todo{Not the best reference}
The hold-out method consists of splitting the observed data set into two
independent parts:
a training set $\mathbf{d}^o_{inv}$
and a testing set $\mathbf{d}^o_{test}$.
The training set is used in the inversion
while the testing set is kept back
and used to judge the quality of the chosen value of $\mu$.
For a value of the regularization parameter $\mu_k$,
the training set is inverted using $\mu_k$
to obtain an estimate $\mathbf{\hat{p}}^k$.
This estimate is used to calculate predicted data
on the same points as the testing set
via forward modeling
($\mathbf{d}_{test}^k = \mathbf{f}(\mathbf{\hat{p}}^k$)).
The metric chosen to evaluate $\mu_k$ is
the mean square error (MSE) of the misfit
between the observed and predicted testing data sets,

\begin{equation}
    MSE_k = \dfrac{\|\mathbf{d}^o_{test} - \mathbf{d}^k_{test}\|^2}{N_{test}},
    \label{eq:msemu}
\end{equation}

\noindent
in which $N_{test}$ is the number of data in the testing set.
The optimal value of $\mu$ will be the one that minimizes the MSE,
i.e. the one that best predicts the testing data.
We emphasize that the inversion is performed only on the training data set.

The algorithm for the hold-out cross-validation is summarized as follows:

\begin{enumerate}
    \item Divide the observed data into
        the training ($\mathbf{d}^o_{inv}$)
        and testing ($\mathbf{d}^o_{test}$) sets.
    \item For each $\mu_k \in [\mu_1, \mu_2, \ldots, \mu_{N_{\mu}}]$:
    \begin{enumerate}
        \item Estimate $\mathbf{\hat{p}}^k$ by inverting the training set
            $\mathbf{d}^o_{inv}$.
        \item Use $\mathbf{\hat{p}}^k$ to calculate the predicted testing set
            $\mathbf{d}^k_{test}$.
        \item Calculate the mean square error $MSE_k$ using Eq.~\ref{eq:msemu}.
    \end{enumerate}
    \item The final solution is the $\mathbf{\hat{p}}^k$ corresponding to the
        smallest $MSE_k$.
\end{enumerate}

The separation of the training and testing data sets is commonly done by taking
random samples from the full data set.
However, we cannot perform the separation in this way because
Bott's method requires data on a regular grid as well as having model elements
directly below each data point.
Thus, we take as our training set the points from the observed data grid that
fall on a similar grid but with twice the grid spacing
(see Fig.~\ref{fig:grid_separation}).
\todo{Include fig from seminar}
All other points from the original data grid
make up the testing data set.
This separation will lead to
a testing data set with more points than the training data set.
A way to balance this loss of data in the inversion
is to generate a data grid with half of the desired grid spacing,
either through interpolation
or from a spherical harmonic model.

\begin{figure}
    \centering
    %\includegraphics[width=\columnwidth]{figures/paper/sam-gravity-data}
    \caption{This is a figure caption.}
    \label{fig:grid_separation}
\end{figure}

\subsection{Estimating $z_{ref}$ and $\Delta\rho$}

The depth of the Normal Earth Moho ($z_{ref}$)
and the density-contrast of the anomalous Moho ($\Delta\rho$)
are other hyper-parameters of the inversion.
That is, their value influences the final solution
but they are not estimated during the inversion.
Both hyper-parameters cannot be determined from the gravity data alone.
Estimating $z_{ref}$ and $\Delta\rho$ requires
information that is independent of the gravity data,
such as knowledge of the parameters at certain points.
This information can be used in a manner similar to
the cross-validation described in the previous section.
In this study, we use point estimates of the Moho depth
to determine the optimal values of $z_{ref}$ and $\Delta\rho$.
These points will generally come from seismologic studies,
like receiver functions, surface wave dispersion, and deep refraction
experiments.

Let $\mathbf{z}_s^o$ be a vector of $N_s$ known Moho depths.
We use the mean square error (MSE)
as a measure of how well a given inversion output $\mathbf{\hat{p}}^k$
fits the know depths.
The optimal values of $z_{ref}$ and $\Delta\rho$
are the ones that best fit the independent known Moho depths
(i.e., produce the smallest MSE).
However, the points do not necessarily coincide
with the model elements of the inversion.
Before computing the MSE,
we interpolate $\mathbf{\hat{p}}^k$ on the known points
to obtain the predicted depths $\mathbf{z}_s^k$.
The MSE is defined as

\begin{equation}
    MSE = \dfrac{\|\mathbf{z}^o_s - \mathbf{z}^k_{s}\|^2}{N_s}.
    \label{eq:msehyper}
\end{equation}

The algorithm for estimating $z_{ref}$ and $\Delta\rho$ is:

\begin{enumerate}
    \item For each combination of
        $z_{ref,l} \in [z_{ref,1},z_{ref,2},\ldots,z_{ref,N_z}]$ and
        $\Delta\rho_m \in
         [\Delta\rho_1,\Delta\rho_2,\ldots,\Delta\rho_{N_{\rho}}]$:
    \begin{enumerate}
        \item Perform the inversion on the training data set
            $\mathbf{d}^o_{inv}$ using $z_{ref,l}$, $\Delta\rho_m$, and
            the previously estimated value of $\mu$.
            The inversion output is the vector $\mathbf{\hat{p}}^{l,m}$.
        \item Interpolate $\mathbf{\hat{p}}^{l,m}$
            on the known points to obtain the predicted depths
            $\mathbf{z}_s^{l,m}$.
        \item Calculate the MSE between $\mathbf{z}_s^o$ and
            $\mathbf{z}_s^{l,m}$ using Eq.~\ref{eq:msehyper}.
    \end{enumerate}
    \item The final solution is the $\mathbf{\hat{p}}^{l,m}$ corresponding to
        the smallest MSE.
\end{enumerate}

A similar approach was used by \citet{martins2010}
to estimate the parameters defining
the density-contrast variation with depth
of a sedimentary basin.
\citet{vandermeijde2013} also had
a similar methodology for dealing with the hyper-parameters,
though in a less formalized way.

\subsection{Software implementation}

The implementation relies on the open-source libraries
scipy and numpy \citep[][ \url{http://scipy.org}]{jones2001}
for array-based computations,
matplotlib \citep[][ \url{http://matplotlib.org}]{hunter2007}
and seaborn
\citep[][ \url{http://stanford.edu/~mwaskom/software/seaborn}]{waskom2015}
for plots and maps,
and Fatiando a Terra \citep[][ \url{http://www.fatiando.org}]{uieda2013}
for geophysics specific tasks (e.g., forward modeling).

The computational experiments
(e.g., data processing, synthetic tests, real data application)
were performed in
Jupyter (formerly IPython) notebooks
\citep[][ \url{http://jupyter.org/}]{perez2007}.
The notebook files combine the source code used to run the experiments,
the results and figures generated by the code,
and rich text to explain and document the analysis.

The Python implementation of the inversion as well as the Jupyter notebooks are
freely available under the terms of the BSD 3-clause open-source license.
Both can be found at the online repository
\url{https://github.com/pinga-lab/paper-moho-inversion-tesseroids}.
An archived version is also available at
\url{http://dx.doi.org/DOI_HERE} (made available upon publication).
\todo{Insert Zenodo DOI here}

\todo{Check which conjugate gradient scipy uses}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Application to synthetic data from a simple model}

Meh.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Application to synthetic data from the CRUST1.0 model}

Meh.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Application to the South American Moho}

\subsection{Gravity data}

The gravitational effect of the topography
is removed using the
ETOPO1 digital terrain model
\citep[][ \url{http://dx.doi.org/10.7289/V5C8276M}]{amante2009}.
The effect of sedimentary basins is removed using the
CRUST1.0 model
\citep[][ \url{http://igppweb.ucsd.edu/~gabi/rem.html}]{laske2013}.
Lateral variations in density along the Moho cannot be properly accounted for
in regions where information coverage is sparse and readily accessible models
are not available, like in the South American and African continents.
For the purposes of this study, we will assume that all other crustal sources
and lateral variations in density are negligible.

%\begin{figure*}
    %\centering
    %\includegraphics[width=\textwidth]{figures/paper/sam-gravity-data}
    %\caption{This is a figure caption.}
    %\label{fig:sam-data}
%\end{figure*}

%\begin{figure*}
    %\centering
    %\includegraphics[width=\textwidth]{figures/paper/sam-gravity-sed}
    %\caption{This is a figure caption.}
    %\label{fig:sam-sed}
%\end{figure*}

\subsection{Inversion results}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}

Meh.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}

Meh.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Acknowledgments}

The authors are indebted to the developers and maintainers of the open-source
software without which this work would not have been possible.

\bibliographystyle{gji}
\bibliography{biblio}

\end{document}