### 2016-02-20 manuscript/profiling.tex
Run synthetic simple profiling and remove steepest
Removed all steepest descent runs from the notebook, leave only the
newton runs. Remove the steepest results files as well. Regenerate
figures from new results. Add profiling results.

\begin{table}
    \centering
    \caption{
        Total time spent on each function during a single inversion of
        simple synthetic data.
        The inversion was performed on a laptop computer with a
        Intel(R) Core(TM) i7-3612QM CPU @ 2.10GHz processor.
        The total time for the inversion was 42.133 seconds.        
    }
    \label{profiling}
    \begin{tabular}{lcc}
        Function description & Time & Percentage of total time\\
        \hline 
        Sparse conjugate gradient & 0.021 s & 0.050\%\\
        Sparse dot product & 0.007 s & 0.017\%\\
        Tesseroid forward modeling & 42.059 s & 99.824\%\\
        \hline
    \end{tabular}
\end{table}