### 2016-03-23 manuscript/profiling.tex
Move units to header of table
Changed this by hand. Need to re-run the notebook later

\begin{table}
    \centering
    \caption{
        Time spent on each function during a single inversion of
        simple synthetic data.
        The inversion was performed on a laptop computer with a
        Intel(R) Core(TM) i7-3612QM CPU @ 2.10GHz processor.
        The total time for the inversion was 42.133 seconds.
    }
    \label{profiling}
    \begin{tabular}{lcc}
        Function description & Time (s) & Percentage of total time (\%)\\
        \hline
        Sparse conjugate gradient & 0.021 & 0.050\\
        Sparse dot product & 0.007 & 0.017\\
        Tesseroid forward modeling & 42.059 & 99.824\\
        \hline
    \end{tabular}
\end{table}