### 2016-08-20 manuscript/manuscript.tex
Minor changes to the text

\documentclass[extra,mreferee]{gji}
%\documentclass[extra]{gji}
\usepackage{timet}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{url}
\usepackage[utf8]{inputenc}
% To make two column figures appear in the correct order
\usepackage{fixltx2e}
% To make annotations on the sides
\usepackage{todonotes}

% Document metadata
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\Title}{
    Fast non-linear gravity inversion in spherical coordinates
    with application to the South American Moho
}
\newcommand{\Keywords}{
    Inverse theory;
    Gravity anomalies and Earth structure;
    Satellite gravity;
    South America;
}
\title[]{\Title}
\author[]{
    Leonardo Uieda$^{1,2}$,
    Valéria C. F. Barbosa$^{2}$
    \\
    $^1$Universidade do Estado do Rio de Janeiro, Rio de Janeiro, Brazil.
    e-mail: leouieda@gmail.com
    \\
    $^2$Observatório Nacional, Rio de Janeiro, Brazil.
}
% Metadata for the generated PDF
\usepackage[pdftex,colorlinks=true]{hyperref}
\hypersetup{
    pdftitle={\Title},
    pdfauthor={Leonardo Uieda (leouieda@gmail.com)},
    pdfsubject={},
    pdfkeywords={\Keywords},
    pdfcreator={pdfTeX},
    allcolors=black,
}

\begin{document}

\label{firstpage}
\maketitle

\begin{abstract}
    Estimating the relief of the Moho from gravity data is a computationally
    intensive non-linear inverse problem.
    What is more, the modeling must take the Earths curvature into account when
    the study area is of regional scale or greater.
    We present a regularized non-linear gravity inversion method that has a low
    computational footprint and employs a spherical Earth approximation.
    To achieve this, we combine the highly efficient Bott's method with
    smoothness regularization and a discretization of the anomalous Moho into
    tesseroids (spherical prisms).
    The computational efficiency of our method is attained by harnessing the
    fact that all matrices involved are sparse.
    The inversion results are controlled by three hyper-parameters: the
    regularization parameter, the anomalous Moho density-contrast, and the
    reference Moho depth.
    We estimate the regularization parameter using the method of hold-out
    cross-validation.
    Additionally, we estimate the density-contrast and the reference depth
    using knowledge of the Moho depth at certain points.
    We apply the proposed method to estimate the Moho depth for the South
    American continent using satellite gravity data and seismological data.
    The final Moho model is in accordance with previous gravity-derived models
    and seismological data.
    The misfit to the gravity and seismological data is worse in the Andes and
    best in oceanic areas, central Brazil and Patagonia, and along the Atlantic
    coast.
    Similarly to previous results, the model suggests a thinner crust of 30-35
    km under the Andean foreland basins.
    Discrepancies with the seismological data are greatest in the Guyana
    Shield, the central Solimões and Amazonas Basins, the Paraná Basin, and the
    Borborema province.
    These differences suggest the existence of crustal or mantle density
    anomalies that were unaccounted for during gravity data processing.
\end{abstract}

\noindent\textbf{Key words:} \Keywords

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

The Mohorovičić discontinuity (or Moho) that marks the transition from the
crust to the mantle, is studied almost exclusively through indirect geophysical
methods.
The two main geophysical methods used to estimate the depth of the Moho are
seismology, with both natural and controlled sources, and gravimetry.
With the advent of satellite gravimetry missions like GRACE and GOCE,
gravity-derived crustal models can be produced in regional or global scales
\citep[e.g. ][]{reguzzoni2013,vandermeijde2013,vandermeijde2015}.
New spherical harmonic gravity models that use these satellite observation,
like GOCO5S \citep{mayer-guerr2015}, provide almost homogeneous data coverage
in difficult to access regions traditionally poor in terrestrial data.
An example is South America, where seismological and terrestrial gravity data
are traditionally concentrated around urban centers and coastal areas,
resulting in large areas (e.g., forests and mountains) devoid of data.

Estimating Moho depth from gravity data is a non-linear inverse problem.
One can generalize this problem of estimating the depths of an interface
separating two media,
such as the sediment-basement interface of a sedimentary basin or the
crust-mantle interface (Moho).
Several methods have been developed over the years to solve this inverse
problem, for example
\citet{bott1960, barbosa1997, barbosa1999b, barbosa1999a, barnes2012, leao1996,
martins2010, martins2011, oldenburg1974, reguzzoni2013, santos2015, silva2006,
silva2014},
to name a few.
Solving the inverse problem is computationally demanding because it requires
the construction of large dense matrices and the solution of large linear
systems.
As a result, some authors search for ways to increase the computational
efficiency of this class of inverse problem.
\citet{bott1960} proposed a method based on iteratively applying corrections to
a starting estimate based on the inversion residuals.
The algorithm is fast because it bypasses the construction and solution of
linear systems and only involves forward modeling.
\citet{oldenburg1974} showed that the fast FFT-based forward modeling of
\citet{parker1973} could be rearranged to estimate the relief.
\citet{barnes2012} use a form of adaptive discretization to compute the
Jacobian, or sensitivity, matrix.
For each data point, the discretization will be progressively coarser
the further way from the point.
This reduces the matrix and, consequently, the linear systems to a sparse form
that can be solved efficiently.
Recently, \citet{silva2014} extended and generalized the original method of
\citet{bott1960} and \citet{santos2015} used this extension to estimate a
basement relief with sharp boundaries.

A spherical Earth approximation is preferred when estimating the Moho depth
from gravity data in continental and global scale studies.
\citet{wieczorek1998} developed a spherical harmonic equivalent of the
Parker-Oldenburg FFT algorithm and applied it to estimate the crustal structure
of the Moon.
\citet{reguzzoni2013} use a spherical Earth approximation to estimate the
global Moho relief using data from the GOCE satellite mission.
Another approach is to use
non-spectral (space domain) gravity inversion methods.
Many such methods were developed for estimating the basement relief of a
sedimentary basin
\citep[e.g., ][]{barbosa1997, barbosa1999a, barbosa1999b, martins2010,
martins2011, sun2014}.
These methods approximate the sedimentary pack by a set of juxtaposed
right-rectangular prisms.
The top of the prisms coincide with the Earth's surface and the prisms'
thicknesses represent the depths to the basement and are the parameters to be
estimated in the inversion.
The use of rectangular prisms implies a planar Earth approximation and may not
be adequate for depth-to-Moho estimates in continental- or global-scale study.
A straightforward way to circumvent this hindrance is to adapt one of the
methods developed for rectangular prisms to use tesseroids (spherical prisms).
One of the difficulties of this approach is that the forward problem for a
tesseroid must be solved numerically.
Two alternatives proposed in the literature to the numerical solution are
Taylor series expansion \citep{heck2007, grombein2013}
and the Gauss-Legendre Quadrature
\citep{asgharzadeh2007}.
Numerical experiments by \citet{wild-pfeiffer2008} suggest that the
Gauss-Legendre Quadrature (GLQ) offers superior results.
However, the GLQ suffers from numerical instability when the computation point
is close to the tesseroid \citep{asgharzadeh2007}.
To overcome the numerical instability, \citet{li2011} proposed an adaptive
discretization algorithm which was later improved upon by \citet{uieda2016}.

In any gravity inversion for estimating the relief of an interface, two
hyper-parameters control the inversion results: the density-contrast between
the two media and the reference level around which the interface undulates.
The reference level is the constant depth of the Normal Earth Moho in the case
of the anomalous Moho.
For regularized inversions, an additional hyper-parameter is the regularization
parameter that balances the relative importance between the data-misfit measure
and the regularizing function.
The two most commonly used methods for estimating the regularization parameter
are the L-curve criterion and Generalized Cross Validation (GCV).
\citet{farquharson2004} provide for a thorough comparison of both methods.
Estimating the density-contrast in a sedimentary basin context has been tackled
by \citet{silva2006} and \citet{martins2010} when the basement depth is known
at a few points.
To the authors' knowledge no attempt has been made to estimate the reference
level.

We present a non-linear gravity inversion to estimate the Moho depth in
a spherical Earth approximation.
Our method is based on the \citet{silva2014} Gauss-Newton formulation of the
method of \citet{bott1960}.
We use tesseroids to discretize the anomalous Moho and the adaptive
discretization algorithm of \citet{uieda2016} for the forward modeling.
The stability of the inversion is achieved through smoothness regularization.
In order to maintain the computational efficiency of Bott's method,
we exploit the sparse nature of all matrices involved in the computations.
We employ a variant of GCV known as hold-out cross-validation \citep{kim2009}
to estimate the regularization parameter.
Additionally, we estimate the density-contrast and reference level
simultaneously in a second validation step using knowledge of the Moho depth at
a few points, similarly to \citet{silva2006} and \citet{martins2010}.
Finally, we apply the proposed method to estimate the Moho depth for South
America using gravity data from the GOCO5S model \citep{mayer-guerr2015} and
the seismological data of \citet{assumpcao2013a}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methodology}

In potential field methods, we must isolate the target anomalous density
distribution before modeling and inversion.
In our case, the target is the relief of the real Moho undulating around a
reference Moho.
We do this by removing all other effects from the gravity observations.
The first correction is to remove the
scalar gravity of an ellipsoidal reference Earth (the Normal Earth),
hereafter denoted as $\gamma$.
This effect is calculated on the same point $P$ where
the gravity observation was made
(Fig~\ref{fig:anomalysketch}a-b).
$\gamma(P)$ is calculated using
the closed-form solution presented by \citet{li2001a}.
The difference between the observed gravity at point P ($g(P)$)
and Normal gravity at the same point
is known as the gravity disturbance,

\begin{equation}
    \delta(P) = g(P) - \gamma(P).
    \label{eq:disturbance}
\end{equation}

The disturbance contains only the gravitational effects of density
distributions that are anomalous with respect to the Normal Earth
(see Fig.~\ref{fig:anomalysketch}c).
This includes all masses above the surface of the ellipsoid (the topography),
the mass deficiency of the oceans,
the mass deficiency of sedimentary basins,
crustal sources (e.g., igneous intrusions, lateral density changes, etc),
heterogeneities below the upper mantle,
and the effect of the difference between the real Moho
topography and the Moho of the Normal Earth.

To estimate the anomalous Moho relief from gravity data,
we must first isolate its gravitational attraction.
Thus, all other gravitational effects
must be either removed or assumed negligible.
Here, we will remove the gravitational effect produced by the known topography
and ocean masses to obtain the full Bouguer disturbance
(Fig~\ref{fig:anomalysketch}d),

\begin{equation}
    \delta_{bg}(P) = \delta(P) - g_{topo}(P).
    \label{eq:bouguer}
\end{equation}

\noindent
We will also remove the gravitational effect of know sedimentary basins
but assume that the effects of
other crustal and mantle sources are negligible.
Thus, the only effect left will be that of the anomalous Moho relief
(Fig~\ref{fig:anomalysketch}e).
The gravitational attraction of the topography, oceans, and basins are
calculated in a spherical Earth approximation by forward modeling using
tesseroids (Fig.~\ref{fig:tesseroid}).
The tesseroid effects are calculated numerically using
Gauss-Legendre Quadrature (GLQ) integration \citep{asgharzadeh2007}.
The accuracy of the GLQ integration is improved by the adaptive discretization
scheme of \citet{uieda2016}.

\begin{figure}
    \centering
    \includegraphics[width=0.5\columnwidth]{figures/problem-concept}
    \caption{
        Sketch of the stages in gravity data correction and
        the discretization of the anomalous Moho relief using tesseroids.
        (a) The Earth and the measured gravity at point P ($g(P)$).
        (b) The Normal Earth and the calculated normal gravity at point P
        ($\gamma(P)$). $z_{ref}$ is the depth of the Normal Earth Moho.
        (c) The gravity disturbance ($\delta(P)$) and the corresponding density
        anomalies after removal of the normal gravity: topography, oceans,
        crustal and mantle heterogeneities, and the anomalous Moho.
        (d) The Bouguer disturbance ($\delta_{bg}(P)$) after topographic
        correction and the remaining density anomalies.
        (e) All density anomalies save the anomalous Moho are assumed to have
        been removed before inversion.
        (f) The discretization of the anomalous Moho in tesseroids. Grey
        tesseroids will have a negative density contrast while red tesseroids
        will have a positive one.
    }
    \label{fig:anomalysketch}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.5\columnwidth]{figures/tesseroid-coord-sys}
    \caption{Sketch of a tesseroid (spherical prism) in a geocentric coordinate
        system (X, Y, Z).
        Observations are made at point P with respect to it's local
        North-oriented coordinate system (x, y, z).
        After \citet{uieda2015}.
    }
    \label{fig:tesseroid}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Parametrization and the forward problem}

We parameterize the forward problem by discretizing the anomalous Moho
into a grid of $M_{lon} \times M_{lat} = M$ juxtaposed tesseroids
(Fig~\ref{fig:anomalysketch}f).
The true (real Earth) Moho varies in depth
with respect to the Moho of the Normal Earth.
Hereafter we will refer to the depth of the Normal Earth Moho as $z_{ref}$
(see Fig.~\ref{fig:anomalysketch}b).
If the true Moho is above $z_{ref}$,
the top of the $k$th tesseroid is the Moho depth $z_{k}$,
the bottom is $z_{ref}$, and the density-contrast ($\Delta\rho$) is positive
(red tesseroids in Fig~\ref{fig:anomalysketch}f).
If the Moho is below $z_{ref}$, the top of the tesseroid is $z_{ref}$,
the bottom is $z_k$, and $\Delta\rho$ is negative
(grey tesseroids in Fig~\ref{fig:anomalysketch}f).

Considering that the absolute value of the density-contrasts
of the tesseroids is a fixed parameter,
the predicted gravity anomaly of the Moho is a non-linear function of the
parameters $z_k$, $k=1, \ldots, M$,

\begin{equation}
    d_i = f_i(\mathbf{p}),
    \label{eq:forward}
\end{equation}

\noindent in which $d_i$ is the $i$th element of the $N$-dimensional predicted
data vector $\mathbf{d}$, $\mathbf{p}$ is the $M$-dimensional parameter vector
containing the $M$ Moho depths ($z_k$),
and $f_i$ is the $i$th non-linear function that maps the parameters onto the
data.
The functions $f_i$ are the radial component of the gravitational attraction
of the tesseroid Moho model.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Inverse problem}

We wish to estimate the parameter vector $\mathbf{p}$ from a set of observed
gravity data $\mathbf{d}^o$.
The least-squares estimate is the one that minimizes the data-misfit function

\begin{equation}
    \phi(\mathbf{p}) =
        [\mathbf{d}^o - \mathbf{d}(\mathbf{p})]^T
        [\mathbf{d}^o - \mathbf{d}(\mathbf{p})].
    \label{eq:data-misfit}
\end{equation}

Function $\phi(\mathbf{p})$ is non-linear with respect to $\mathbf{p}$.
Thus, we can determine its minimum using gradient-based
iterative optimization
methods like Gauss-Newton or Steepest Descent.
Such methods start from an initial approximation to the model parameter vector
$\mathbf{p}^0$ and estimate a parameter perturbation vector
$\mathbf{\Delta p}^0$.
The perturbation vector is used to update $\mathbf{p}^0$ to
$\mathbf{p}^1 =  \mathbf{p}^0 + \mathbf{\Delta p}^0$.
This procedure is repeated until a minimum of function $\phi(\mathbf{p})$
(Eq.~\ref{eq:data-misfit}) is reached.

For the Gauss-Newton method,
the parameter perturbation vector at the $k$th iteration $\mathbf{\Delta p}^k$
is obtained by solving the linear system

\begin{equation}
    \mathbf{H}^k\mathbf{\Delta p}^k = -\mathbf{\nabla\phi}^k,
    \label{eq:gaussnewton}
\end{equation}

\noindent in which
$\mathbf{\nabla\phi}^k$ and $\mathbf{H}^k$ are, respectively,
the gradient vector and the Hessian matrix of $\phi(\mathbf{p})$.

The gradient vector and the Gauss-Newton approximation of the Hessian matrix
of $\phi(\mathbf{p})$ are, respectively,

\begin{equation}
    \mathbf{\nabla\phi}^k = -2{\mathbf{A}^k}^T[\mathbf{d}^o - \mathbf{d}(\mathbf{p}^k)],
    \label{eq:gradient}
\end{equation}

\noindent
and

\begin{equation}
    \mathbf{H}^k \approx 2{\mathbf{A}^k}^T{\mathbf{A}^k},
    \label{eq:hessian}
\end{equation}

\noindent in which
$\mathbf{A}^k$ is the $N \times M$ Jacobian or sensitivity matrix
whose elements are

\begin{equation}
    A_{ij}^k = \dfrac{\partial f_i}{\partial p_j}(\mathbf{p}^k).
    \label{eq:jacobian}
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Regularization}

Non-linear gravity inversions for estimating the relief of an interface
separating two media (like the Moho) are ill-posed and require additional
constraints in the form of regularization \citep{silva2001b}.
A common approach is to use the first-order Tikhonov regularization
\citep{tikhonov1977} to impose smoothness on the solution.
The cost function for smoothness regularization is given by

\begin{equation}
    \theta(\mathbf{p}) = \mathbf{p}^T\mathbf{R}^T\mathbf{R}\mathbf{p},
    \label{eq:regul}
\end{equation}

\noindent where $\mathbf{R}$ is an $L \times M$ finite-difference matrix
representing $L$ first-order differences between the depths of adjacent
tesseroids.

To transform the ill-posed inverse problem into a well-posed one via Tikhonov
regularization, we adopted the well-established procedure
of formulating a constrained inverse problem that is solved by minimizing an
unconstrained goal function

\begin{equation}
    \Gamma(\mathbf{p}) = \phi(\mathbf{p}) + \mu\theta(\mathbf{p}),
    \label{eq:goalfunction}
\end{equation}

\noindent
in which $\mu$ is the regularization parameter that controls the balance
between fitting the observed data and obeying the smoothness constraint imposed
by the regularizing function $\theta(\mathbf{p})$ (Eq.~\ref{eq:regul}).

The goal function $\Gamma(\mathbf{p})$ is also non-linear with respect to
$\mathbf{p}$ and can be minimized using the Gauss-Newton method.
The gradient vector and Hessian matrix of the goal function are, respectively,

\begin{equation}
    \mathbf{\nabla\Gamma}^k =
        -2{\mathbf{A}^k}^T[\mathbf{d}^o - \mathbf{d}(\mathbf{p}^k)] +
        2\mu\mathbf{R}^T\mathbf{R}\mathbf{p}^k,
    \label{eq:gradient-regul}
\end{equation}

\noindent and

\begin{equation}
    \mathbf{H}^k = 2{\mathbf{A}^k}^T{\mathbf{A}^k}
                   + 2\mu\mathbf{R}^T\mathbf{R}.
    \label{eq:hessian-regul}
\end{equation}

At the $k$th iteration, the parameter perturbation vector
$\mathbf{\Delta p}^k$ is obtained by solving the linear equation system

\begin{equation}
    \left[{\mathbf{A}^k}^T{\mathbf{A}^k} + \mu\mathbf{R}^T\mathbf{R}\right]
    \mathbf{\Delta p}^k =
    {\mathbf{A}^k}^T[\mathbf{d}^o - \mathbf{d}(\mathbf{p}^k)] -
    \mu\mathbf{R}^T\mathbf{R}\mathbf{p}^k.
    \label{eq:gaussnewton-regul}
\end{equation}

Estimating the Moho depths using the above equations is computationally
costly because of two main factors:
(1) the evaluation and storage of the dense $N \times M$ Jacobian matrix
${\mathbf{A}^k}$
and (2) the solution of the resulting $M \times M$ equation system.
In practice, the derivatives in the Jacobian (Eq.~\ref{eq:jacobian})
are often calculated through a first-order finite-difference approximation.
Thus, evaluating ${\mathbf{A}^k}$ requires $2\times M \times N$ forward modeling
operations for each iteration of the gradient descent algorithm.
These computations are performed for each iteration of the optimization
of the goal function $\Gamma(\mathbf{p})$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Bott's method}

\citet{bott1960} developed an efficient method to determined the depth of the
basement of a sedimentary basin from gravity observations.
The method requires data on a regular grid of $N_x \times N_y = N$
observations.
The basement relief is then discretized into an equal grid of $M_x \times
M_y = M$ elements with $M_x = N_x$ and $M_y = N_y$.
Bott's iterative method starts with an initial approximation of the basement
depths $\mathbf{p}^0$ equal to the null vector.
The method updates the approximation by calculating a parameter perturbation
vector $\mathbf{\Delta p}^k$ using the formula

\begin{equation}
    \mathbf{\Delta p}^k =
        \dfrac{\mathbf{d}^o - \mathbf{d}(\mathbf{p}^k)}{2\pi G \Delta \rho},
    \label{eq:bott}
\end{equation}

\noindent
in which $G$ is the gravitational constant and $\Delta \rho$ is the
contrast between the density of the sediments and the reference density.
The iterative process stops when the inversion residuals
$\mathbf{r}^k = \mathbf{d}^o - \mathbf{d}(\mathbf{p}^k)$ fall below the assumed noise level
of the data.

\citet{silva2014} showed that Bott's method can be formulated as
a special case of the Gauss-Newton method (Eq.~\ref{eq:gaussnewton})
by setting the Jacobian matrix (Eq.~\ref{eq:jacobian}) to

\begin{equation}
    \mathbf{A} = 2\pi G \Delta \rho \mathbf{I},
    \label{eq:bott-gaussnewton}
\end{equation}

\noindent
where $\mathbf{I}$ is the identity matrix.
In this framework,
Bott's method uses a Bouguer plate approximation of the gravitational effect of
the relief, $d_i = 2\pi G \Delta\rho\ p_i$.
The derivative of $d_i$ with respect to the parameter $p_i$ is
$2\pi G \Delta \rho$, thus linearizing the Jacobian matrix.
However, the non-linearity of the predicted data $\mathbf{d}(\mathbf{p}^k)$ is
preserved.

One of the advantages of Bott's method over the traditional Gauss-Newton or
Steepest Descent is the elimination of the computation and storage of the dense
Jacobian matrix $\mathbf{A}^k$.
Furthermore, Bott's method also does not require the solution of equation
systems.
However, a disadvantage of Bott's method is that it suffers from instability
\citep{silva2014}.
A common approach to counter this issue is to apply a smoothing filter after
the inversion to the unstable estimate, as in \citet{silva2014}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Combining Bott's method, regularization, and tesseroids}

We propose a regularized version of Bott's method to invert gravity data for
estimating the depth of the Moho in spherical coordinates.
To adapt Bott's method to spherical coordinates,
we replace the right-rectangular prisms in the forward modeling
($\mathbf{d}(\mathbf{p}^k)$ in Eq.~\ref{eq:bott})
with tesseroids.
The tesseroid forward modeling uses the adaptive discretization algorithm
of \citet{uieda2016} to achieve accurate results.
Furthermore, our formulation maintains the regularized solution
for the Gauss-Newton method (Eq.~\ref{eq:gaussnewton-regul})
but replaces the full Jacobian matrix with the Bouguer plate approximation.
Here, the Jacobian matrix is replaced by a diagonal matrix
(Eq.~\ref{eq:bott-gaussnewton}) whose elements are invariant along successive
iterations.
Using this approximation eliminates the cost of computing and storing
the full $N \times M$-dimensional Jacobian matrix $\mathbf{A}^k$ at each
iteration
(Eq.~\ref{eq:jacobian}).
Traditionally, the full Jacobian matrix is computed using a first-order finite
difference scheme,
which requires $2 \times N \times M$ forward modeling operations per iteration.
Using Eq.~\ref{eq:bott-gaussnewton} requires $N$ multiplications that need only
be performed once.
This provides a considerable speed gain.

Matrix arithmetic operations can be performed efficiently by taking advantage
of the sparse nature of matrices $\mathbf{A}$ and $\mathbf{R}$
(respectively, Eq.~\ref{eq:bott-gaussnewton} and \ref{eq:regul}).
The same is true for solving the equation system in the Gauss-Newton method
(Eq.~\ref{eq:gaussnewton-regul}).
However, the computational cost of forward modeling is still present.
Particularly, forward modeling using tesseroids is more computationally
intensive than using right-rectangular prisms
because of the numerical integration and adaptive discretization
\citep{uieda2016}.
We show later in this article that sparse matrix multiplications and solving
the sparse linear system in Eq.~\ref{eq:gaussnewton-regul} account for less
than 0.1\% of the computation time required for a single inversion.

The main advantage of our formulation is that it retains the
efficiency of Bott's method while stabilizing the solution through the well
established formalism of Tikhonov regularization.
For example, the total variation approach used by \citet{martins2011} could
potentially be implemented in a more straight forward manner than done by
\citet{santos2015}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Estimating the inversion hyper-parameters}

Parameters that influence the inversion result but are not estimated directly
in the inversion are known as hyper-parameters.
In the case of our regularized Moho depth inversion, the hyper-parameters are
the regularization parameter $\mu$ (Eq.~\ref{eq:goalfunction}),
the Moho density-contrast $\Delta \rho$ (Eq.~\ref{eq:bott-gaussnewton}),
and the depth of the Normal Earth Moho, or reference level, $z_{ref}$
(Fig.~\ref{fig:anomalysketch}b).

We estimate these hyper-parameters in two steps.
First, we assume fixed values for
$z_{ref}$ and $\Delta\rho$ and perform a hold-out cross-validation procedure
\citep{hansen1992} to estimate an optimal value for $\mu$.
Our investigations suggest that the optimal value of $\mu$ does not depend on
the particular values of $z_{ref}$ and $\Delta\rho$ used.
Second, we use the estimated $\mu$ to perform a validation
procedure to estimate $z_{ref}$ and $\Delta\rho$.
The final outcomes of both steps are the values of the three hyper-parameters
and the final estimated Moho depths.

\subsubsection{Estimating the regularization parameter}

\begin{figure}
    \centering
    \includegraphics[width=0.5\columnwidth]{figures/cv-grid-separation}
    \caption{Sketch of a data grid separated into
        the training (open circles)
        and testing (black dots) data sets.
        The training data set is still displayed on a regular grid
        but with twice the grid spacing
        of the original data grid.}
    \label{fig:grid_separation}
\end{figure}

The regularization parameter $\mu$ controls how much smoothness is applied to
the inversion result.
An optimal value of $\mu$ will stabilize and smooth the solution while not
compromising the fit to the observed data.
Two widely used methods to estimate an optimal $\mu$ are
the L-curve criterion and cross-validation \citep{hansen1992}.
Here, we will adopt the hold-out method of cross-validation \citep{kim2009}.
The hold-out method consists of splitting the observed data set into two
independent parts:
a training set $\mathbf{d}^o_{inv}$
and a testing set $\mathbf{d}^o_{test}$.
The training set is used in the inversion
while the testing set is kept back
and used to judge the quality of the chosen value of $\mu$.
For a value of the regularization parameter $\mu_n$,
the training set is inverted using $\mu_n$
to obtain an estimate $\mathbf{\hat{p}}^n$.
This estimate is used to calculate predicted data
on the same points as the testing set
via forward modeling

\begin{equation}
    \mathbf{d}_{test}^n = \mathbf{f}(\mathbf{\hat{p}}^n).
    \label{eq:predicted_testing}
\end{equation}

The metric chosen to evaluate $\mu_n$ is
the mean square error (MSE) of the misfit
between the observed and predicted testing data sets,

\begin{equation}
    MSE_n = \dfrac{\|\mathbf{d}^o_{test} - \mathbf{d}^n_{test}\|^2}{N_{test}},
    \label{eq:msemu}
\end{equation}

\noindent
in which $N_{test}$ is the number of data in the testing set.
The optimal value of $\mu$ will be the one that minimizes the MSE,
i.e. the one that best predicts the testing data.
We emphasize that the inversion is performed on the training data set only.

The algorithm for the hold-out cross-validation is summarized as follows:

\begin{enumerate}
    \item Divide the observed data into
        the training ($\mathbf{d}^o_{inv}$)
        and testing ($\mathbf{d}^o_{test}$) sets.
    \item For each $\mu_n \in [\mu_1, \mu_2, \ldots, \mu_{N_{\mu}}]$:
    \begin{enumerate}
        \item Estimate $\mathbf{\hat{p}}^n$ by inverting the training set
            $\mathbf{d}^o_{inv}$.
        \item Use $\mathbf{\hat{p}}^n$ to calculate the predicted testing set
            $\mathbf{d}^n_{test}$ using Eq.~\ref{eq:predicted_testing}.
        \item Calculate the mean square error $MSE_n$ using Eq.~\ref{eq:msemu}.
    \end{enumerate}
    \item The final solution is the $\mathbf{\hat{p}}^n$ corresponding to the
        smallest $MSE_n$.
\end{enumerate}

The separation of the training and testing data sets is commonly done by taking
random samples from the full data set.
However, we cannot perform the separation in this way because
Bott's method requires data on a regular grid as well as having model elements
directly below each data point.
Thus, we take as our training set the points from the observed data grid that
fall on a similar grid but with twice the grid spacing
(open circles in Fig.~\ref{fig:grid_separation}).
All other points from the original data grid
make up the testing data set
(black dots in Fig.~\ref{fig:grid_separation}).
This separation will lead to
a testing data set with more points than the training data set.
A way to balance this loss of data in the inversion
is to generate a data grid with half of the desired grid spacing,
either through interpolation
or from a spherical harmonic model.

\subsubsection{Estimating $z_{ref}$ and $\Delta\rho$}

The depth of the Normal Earth Moho ($z_{ref}$)
and the density-contrast of the anomalous Moho ($\Delta\rho$)
are other hyper-parameters of the inversion.
That is, their value influences the final solution
but they are not estimated during the inversion.
Both hyper-parameters cannot be determined from the gravity data alone.
Estimating $z_{ref}$ and $\Delta\rho$ requires
information that is independent of the gravity data,
such as knowledge of the parameters (Moho depths) at certain points.
This information can be used in a manner similar to
the cross-validation described in the previous section.
In this study, we use point estimates of the Moho depth
to determine the optimal values of $z_{ref}$ and $\Delta\rho$.
These points will generally come from seismologic studies,
like receiver functions, surface wave dispersion, and deep refraction
experiments.

Let $\mathbf{z}_s^o$ be a vector of $N_s$ known Moho depths.
We use the mean square error (MSE)
as a measure of how well a given inversion output $\mathbf{\hat{p}}^{l,m}$
fits the know depths.
The optimal values of $z_{ref}$ and $\Delta\rho$
are the ones that best fit the independent known Moho depths
(i.e., produce the smallest MSE).
However, the points do not necessarily coincide
with the model elements of the inversion.
Before computing the MSE,
we interpolate $\mathbf{\hat{p}}^{l,m}$ on the known points
to obtain the predicted depths $\mathbf{z}_s^{l,m}$.
The MSE is defined as

\begin{equation}
    MSE = \dfrac{\|\mathbf{z}^o_s - \mathbf{z}^{l,m}_{s}\|^2}{N_s}.
    \label{eq:msehyper}
\end{equation}

The algorithm for estimating $z_{ref}$ and $\Delta\rho$ is:

\begin{enumerate}
    \item For every combination of
        $z_{ref,l} \in [z_{ref,1},z_{ref,2},\ldots,z_{ref,N_z}]$ and
        $\Delta\rho_m \in
         [\Delta\rho_1,\Delta\rho_2,\ldots,\Delta\rho_{N_{\rho}}]$:
    \begin{enumerate}
        \item Perform the inversion on the training data set
            $\mathbf{d}^o_{inv}$ using $z_{ref,l}$, $\Delta\rho_m$, and
            the previously estimated value of $\mu$.
            The inversion output is the vector $\mathbf{\hat{p}}^{l,m}$.
        \item Interpolate $\mathbf{\hat{p}}^{l,m}$
            on the known points to obtain the predicted depths
            $\mathbf{z}_s^{l,m}$.
        \item Calculate the MSE between $\mathbf{z}_s^o$ and
            $\mathbf{z}_s^{l,m}$ using Eq.~\ref{eq:msehyper}.
    \end{enumerate}
    \item The final solution is the $\mathbf{\hat{p}}^{l,m}$ corresponding to
        the smallest MSE.
\end{enumerate}

A similar approach was used by \citet{silva2006} and \citet{martins2010}
to estimate the parameters defining
the density-contrast variation with depth
of a sedimentary basin.
\citet{vandermeijde2013} also had
a similar methodology for dealing with the hyper-parameters,
though in a less formalized way.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Software implementation}\label{sec:software}

The inversion method proposed here is implemented in the Python programming
language.
The software is freely available
under the terms of the BSD 3-clause open-source software license.
Our implementation relies on the open-source libraries
scipy and numpy \citep[][ \url{http://scipy.org}]{jones2001}
for array-based computations,
matplotlib \citep[][ \url{http://matplotlib.org}]{hunter2007}
and seaborn
\citep[][ \url{http://stanford.edu/~mwaskom/software/seaborn}]{waskom2015}
for plots and maps,
and Fatiando a Terra \citep[][ \url{http://www.fatiando.org}]{uieda2013a}
for geophysics specific tasks,
particularly for forward modeling using tesseroids.
We use the scipy.sparse package for sparse matrix arithmetic and linear
algebra.
The sparse linear system in Eq.~\ref{eq:gaussnewton-regul}
is solved using the conjugate gradient method implemented in scipy.sparse.

The computational experiments
(e.g., data processing, synthetic tests, real data application)
were performed in
Jupyter (formerly IPython) notebooks
\citep[][ \url{http://jupyter.org/}]{perez2007}.
The notebook files combine the source code used to run the experiments,
the results and figures generated by the code,
and rich text to explain and document the analysis.

All source code, Jupyter notebooks, data, and results
can be found at the online repository
\url{https://github.com/pinga-lab/paper-moho-inversion-tesseroids}.
The repository also contains instructions for replicating all results presented
here.
An archived version of this repository is also available at
\url{http://dx.doi.org/...}
(\textbf{Note to reviewers: the archived version will be uploaded upon
publication}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Application to synthetic data}

We test and illustrate the proposed inversion method
by applying it to two noise-corrupted synthetic data sets.
The first one is generated by a simple Moho model simulating the transition
from a thicker continental crust to a thinner oceanic crust.
This application uses cross-validation to estimate the regularizing parameter
($\mu$) while assuming that the anomalous Moho density-contrast ($\Delta\rho$)
and the Normal Earth Moho depth ($z_{ref}$) are known quantities.
This first test is simplified in order to investigate solely
the efficiency of the inversion and
the cross-validation procedure to estimate $\mu$.
The second data set is generated by a more complex model derived from
the South American portion of the global CRUST1.0 model \citep{laske2013}.
This second application uses cross-validation to estimate $\mu$ and
the validation procedure using synthetic seismological data
to estimate $\Delta\rho$ and $z_{ref}$.
The model and corresponding synthetic data are meant to simulate
with more fidelity the real data application.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Simple model}\label{sec:simple-synthetic}

\begin{figure}
    \centering
    \includegraphics[width=0.5\columnwidth]{figures/synthetic-simple-data}
    \caption{
        A simple Moho model made of tesseroids for synthetic data application.
        (a) The Moho depth of the model in kilometers.
        The model transitions from a deep Moho in the right to a shallow Moho in
        left, simulating the transition between a continental and an oceanic
        Moho.
        Each pixel in the pseudo-color image corresponds to a tesseroid of the
        model.
        (b) Noise-corrupted synthetic gravity data generated from the model
        shown in (a).
    }
    \label{fig:simple-data}
\end{figure}

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figures/synthetic-simple-results}
    \caption{
        Results from the inversion of the simple synthetic data.
        (a) The estimated Moho depth.
        (b) The Moho depth residuals (difference between the true and estimated
        Moho depths).
        (c) The gravity residuals (difference between the observed and
        predicted gravity data).
        (d) Histogram of the gravity residuals shown in c, with the calculated
        mean and standard deviation (std) of the residuals in mGal.
        (e) Cross-validation curve used to determine the optimal regularization
        parameter (Eq.~\ref{eq:goalfunction}).
        Both axis are in logarithmic scale.
        The minimum Mean Square Error (Eq.~\ref{eq:msemu}) is found at
        $\mu = 0.00046$ (red triangle).
        (f) Goal function value (Eq.~\ref{eq:goalfunction}) per Gauss-Newton
        iteration showing the convergence of the gradient descent.
        The y-axis is in logarithmic scale.
    }
    \label{fig:simple-results}
\end{figure*}

We simulate the transition from a continental-type Moho to an oceanic-type Moho
using a model composed of $M_{lat} \times M_{lon} = 40 \times 50$ grid of
juxtaposed tesseroids (a total of $M = 2000$ model elements).
The anomalous Moho density-contrast is $\Delta\rho = 400\ kg/m^3$
and the Normal Earth Moho depth is $z_{ref} = 30\ km$.
Fig.~\ref{fig:simple-data}a shows the model Moho depths where we can clearly
see an eastward crustal thinning.
In Fig.~\ref{fig:simple-data}a, each pixel in the pseudo-color image
corresponds to a tesseroid of the model.

The synthetic data were forward modeled on a regular grid of
$N_{lat} \times N_{lon} = 79 \times 99$ points
(a total of $N = 7821$ observations)
at a constant height of 50 km.
The data were contaminated with pseudo-random noise
sampled from a normal distribution with zero mean and 5 mGal standard deviation.
Fig.~\ref{fig:simple-data}b shows the noise-corrupted full synthetic data set
exhibiting an eastward increase due to the simulated eastward
crustal thinning shown in Fig.~\ref{fig:simple-data}a.
The data grid spacing is half the grid spacing of the tesseroid model
so that, when separating the training and testing data sets
(Fig.~\ref{fig:grid_separation}),
the training data set points will fall directly above each model element.

We separated the synthetic data into training and testing data sets
following Fig.~\ref{fig:grid_separation}.
The training data set is a regular grid of
$N_{lat} \times N_{lon} = 40 \times 50$ points
(a total of $N_{train} = 2000$).
The testing data set is composed of $N_{test} = 5821$ observations.
We used cross-validation to estimate an optimal regularization parameter ($\mu$)
from a set of $N_\mu = 16$ values equally spaced on a logarithmic scale
between $10^{-6}$ and $10^{-1}$.
We ran our regularized inversion on the training data set
for each value of $\mu$,
obtaining 16 Moho depth estimates.
For all inversions, the initial Moho depth estimate
used to start the Gauss-Newton optimization
was set to 60 km depth for all inversion parameters.
Furthermore, $z_{ref}$ and $\Delta\rho$ are set to their respective true values.
Finally, we computed the mean square error (MSE, Eq.~\ref{eq:msemu})
for each estimate and chose as the final estimated Moho model
the one that minimizes the MSE.

Fig.~\ref{fig:simple-results}a shows the final estimated Moho depth
after the cross-validation.
The recovered model is smooth, indicating that the cross-validation procedure
was effective in estimating an optimal regularization parameter.
Fig.~\ref{fig:simple-results}b shows difference between the true Moho depth
(Fig.~\ref{fig:simple-data}a) and the estimated Moho depth.
The differences appear to be semi-randomly distributed with a maximum
coinciding with a short-wavelength feature in the true model.
The maximum and minimum differences are approximately
2.19 and -2.13 km, respectively.
Fig.~\ref{fig:simple-results}c shows inversion residuals, defined as the
difference between the observed and predicted data (in mGal).
The largest residual (in absolute value) coincides with the largest difference
between the true model and the estimate.
The inversion residuals are normally distributed,
as shown in Fig.~\ref{fig:simple-results}d,
with 0.02 mGal mean and a standard deviation of 3.63 mGal.
The cross-validation curve in Fig.~\ref{fig:simple-results}e
shows a clear minimum MSE at $\mu = 0.00046$
(indicated by the red triangle).
Fig.~\ref{fig:simple-results}f shows the convergence of
the Gauss-Newton optimization in eight iterations.

We also investigated the computation time spent in each section of the inversion
process using a source code profiler.
The profiler measures how much time is spent inside each function during
the execution of a program.
We ran the profiler on a single inversion of the training data set
using the estimated regularization parameter.
We tracked the total time spent inside each of the three functions
that represent the potential bottlenecks of the inversion:
solving the linear system in Eq.~\ref{eq:gaussnewton-regul}
using the conjugate gradient method,
performing the dot products required to compute
the Hessian matrix (Eq.~\ref{eq:hessian-regul})
and the gradient vector (Eq.~\ref{eq:gradient-regul}),
and forward modeling to calculate the predicted data (Eq.~\ref{eq:forward}).
The profiling results presented in Table~\ref{profiling}
show that the time spent on forward modeling accounts for approximately
99.8\% of the total computation time.

% Insert the profiling results table
\input{profiling.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Model based on CRUST1.0}\label{sec:crust1}

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figures/synthetic-crust1-data}
    \caption{
        Synthetic data of a model derived from CRUST1.0.
        The model is made of tesseroids with an constant density-contrast
        of $\Delta\rho = 350\ kg/m^3$ and assuming a reference level of
        $z_{ref} = 30\ km$.
        (a) The Moho depth of the model in kilometers.
        Each pixel in the pseudo-color image corresponds to a tesseroid of the
        model.
        (b) Noise-corrupted synthetic gravity data generated from the model.
        (c) Simulated points where the Moho depths are known from seismological
        estimates (color dots).  Here, these point were obtained by
        interpolating the Moho depth in (a).
    }
    \label{fig:crust1-data}
\end{figure*}

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figures/synthetic-crust1-results}
    \caption{
        Inversion results from the CRUST1.0 synthetic data.
        (a) Cross-validation curve used to determine
        the regularization parameter (Eq.~\ref{eq:goalfunction}).
        The minimum Mean Square Error (Eq.~\ref{eq:msemu}) is found at
        $\mu = 0.0001$ (red triangle).
        (b) Validation results used to determine
        the reference level ($z_{ref}$) and the density-contrast ($\Delta\rho$).
        The colors represent the Mean Square Error (Eq.~\ref{eq:msehyper}) in
        $km^2$.
        The minimum (red triangle) is found at $z_{ref} = 30\ km$
        and $\Delta\rho = 350\ kg/m^3$.
        (c) The estimated Moho depth.
        (d) Difference between the CRUST1.0 model depths
        (Fig.~\ref{fig:crust1-data}a)
        and the estimated depths.
        (e) Histogram of the inversion residuals
        (observed minus predicted data).
        (f) Histogram of the differences between
        the synthetic seismic observations (Fig.~\ref{fig:crust1-data}c)
        and the estimated depths.
        (g) The inversion residuals.
        (h) Difference between the seismic and the estimated depths.
    }
    \label{fig:crust1-results}
\end{figure*}

In this test, we simulate the anomalous Moho of South America
using Moho depth information extracted from the CRUST1.0 model
\citep{laske2013}.
We construct a tesseroid model with
$M_{lat} \times M_{lon} = 80 \times 60$ juxtaposed elements, 4800 in total,
using the Moho depths shown in Fig.~\ref{fig:crust1-data}a.
In our model, the Normal Earth Moho is $z_{ref} = 30\ km$ and
the density-contrast is $\Delta\rho = 350\ kg/m^3$.
We produce the synthetic data at a constant height of 50 km
and on a regular grid of $N_{lat} \times N_{lon} = 159 \times 119$ points
(a total of 18,921 observations).
We contaminate the synthetic data with normally distributed pseudo-random noise
with zero mean and 5 mGal standard deviation (Fig.~\ref{fig:crust1-data}b).

The validation procedure to determine $\Delta\rho$ and $z_{ref}$
requires knowledge of the Moho depth at certain points
($\mathbf{z}_s^o$ in Eq.~\ref{eq:msehyper}),
usually from seismic experiments.
Thus, we must also generate synthetic seismic data about the Moho depth.
We produce such data by interpolating the Moho depth shown in
Fig.~\ref{fig:crust1-data}a on the same 937 geographic coordinates
pinpointed in the data set of \citet{assumpcao2013a}.
The resulting synthetic seismic data is shown in Fig.~\ref{fig:crust1-data}c.

We estimate the three hyper-parameters in two parts.
First, we run the cross-validation to estimate
an optimal regularization parameter ($\mu$).
The starting estimate for all inversions is
60 km depth for all model parameters.
For this cross-validation,
we keep $z_{ref}$ and $\Delta\rho$ fixed to
$20\ km$ and $500\ kg/m^3$, respectively.
Second, we use the estimated $\mu$ to run the validation
procedure with the synthetic seismological data
to estimate $z_{ref}$ and $\Delta\rho$,
thus obtaining the final estimated Moho depths.
Fig.~\ref{fig:crust1-results} summarizes the results.

For the cross-validation,
we separate the synthetic data (Fig.~\ref{fig:grid_separation}) into
a training set with twice the grid spacing of the original data
($N_{lat} \times N_{lon} = 80 \times 60$)
and a testing set with 14,121 observations.
We run the inversion for 16 different values of $\mu$
equally spaced in a logarithmic scale between $10^{-7}$ and $10^{-2}$.
For each of the 16 estimates we compute the MSE (Eq.~\ref{eq:msemu}),
shown in Fig.~\ref{fig:crust1-results}a as function of $\mu$.
The optimal regularization parameter that minimizes the MSE is $\mu = 10^{-4}$
(red triangle in Fig.~\ref{fig:crust1-results}a).

In the validation using seismological data,
we use the estimated value of $\mu$ in all inversions.
We test seven values of $z_{ref}$ from 20 to 35 km with 2.5 km intervals
and seven values of $\Delta\rho$ from 200 to 500 $kg/m^3$
with 50 $kg/m^3$ intervals.
We run the inversion for every combination of $z_{ref}$ and $\Delta\rho$,
totaling 49 inversions.
Finally, we calculate the Mean Square Error (Eq.~\ref{eq:msehyper})
for each of the 49 estimates
and choose the values of $z_{ref}$ and $\Delta\rho$ that minimize the MSE.
Fig.~\ref{fig:crust1-results}b
shows a pseudo-color map of the MSE
with a minimum (marked by the red triangle)
at $z_{ref} = 30\ km$ and $\Delta\rho = 350\ kg/m^3$.

Fig.~\ref{fig:crust1-results}c shows the final solution after
cross-validation and validation using seismological data.
The recovered model is smooth, indicating that the cross-validation procedure
was effective in estimating an optimal regularization parameter.
Fig.~\ref{fig:crust1-results}d shows the difference between
the true Moho depths (Fig.~\ref{fig:crust1-data}a)
and the estimated depths (Fig.~\ref{fig:crust1-results}c).
The maximum and minimum differences are, respectively,
9.8 and -8.2 km.
The largest absolute differences are located along the central and northern
Andes, where there is a sharp increase in the true Moho depth
(Fig.~\ref{fig:crust1-data}a).
Positive differences (indicating a too shallow estimate)
appear along the central portion of the Andes,
flanked by regions of negative differences (indicating a too deep estimate)
on the continental and Pacific sides.
Figs.~\ref{fig:crust1-results}e and g show the gravity residuals,
defined as the difference between the observed and predicted gravity data.
The residuals appear normally distributed,
with 0.03 mGal mean and a standard deviation of 4.10 mGal.
The gravity residuals follow a similar, though reversed, pattern
to the differences shown in Fig.~\ref{fig:crust1-results}d.
The largest residuals (in absolute value) are along the Andes,
with the central portion being dominated by negative residuals
and flanked by positive residuals on both sides.
Figs.~\ref{fig:crust1-results}f and h show the differences between
the synthetic seismic data (Fig.~\ref{fig:crust1-data}c)
and the estimated Moho depths.
Once more, the largest differences are concentrated along the Andes,
particularly in the central Andes and near Ecuador and Colombia.
The differences are smaller along the Atlantic coast of South America,
with notable larger differences in a few points of northeastern Brazil
and along the Amazon river.
In general, large residuals are associated with sharp increases in Moho depth.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Application to the South American Moho}

We apply the inversion method proposed here to invert for the Moho depth of the
South American continent.
We follow the application of \citet{vandermeijde2013} but with some
differences, mainly using a different data set and performing all modeling
in spherical coordinates using tesseroids.
The data are corrected of the effects of topography and sedimentary basins.
Crust and mantle heterogeneities cannot be properly accounted for
in regions where information coverage is sparse and readily accessible models
are not available, like in South America and Africa.
Hence, for the purposes of this study, we will assume to be negligible all
other crustal and mantle sources, including lateral variations in density along
the Moho.
All tesseroid models are defined with respect to a spherical Earth of radius
6,378.137 km.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Gravity and seismological data}

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figures/south-america-corrections}
    \caption{
        Gravity data for South America and the models used in the data
        corrections.
        (a) The gravity disturbance (Eq.~\ref{eq:disturbance}) calculated from
        the raw gravity data.
        (b) Topography from ETOPO1.
        (c) Gravitational attraction of the topography calculated
        at the observation height using tesseroids.
        (d) The Bouguer disturbance (Eq.~\ref{eq:bouguer}) obtained by
        subtracting (c) from (a).
        The upper (e), middle (f), and lower (g) sediment layer thicknesses
        from the CRUST1.0 model.
        (h) The total gravitational attraction of the sediment layers shown in
        (e), (f), and (g), calculated using tesseroids.
        }
    \label{fig:sam-corrections}
\end{figure*}

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figures/south-america-data}
    \caption{
        Input data for the South American Moho inversion.
        (a) Sediment-free Bouguer disturbance for South America.
        Obtained by subtracting the total sediment gravitational effect
        (Fig.~\ref{fig:sam-corrections}h) from the Bouguer disturbance
        (Fig.~\ref{fig:sam-corrections}d).
        (b) Seismological Moho depth estimates from
        \citet{assumpcao2013a}.
    }
    \label{fig:sam-data}
\end{figure*}

The raw gravity data are generated from the satellite only
spherical harmonic model GOCO5S \citet{mayer-guerr2015}.
The GOCO5S model combines data from 15 satellites, including the complete
mission data from the GOCE satellite.
The data were downloaded from the
International Centre for Global Earth Models (ICGEM) web-service
\citep[][ \url{http://icgem.gfz-potsdam.de/ICGEM/}]{barthelmes2012}
in the form of the complete gravity field
on a regular grid with $0.2^\circ$ grid spacing at ellipsoidal height 50 km.
We calculate the gravity disturbance
($\delta(P)$ in Eq.~\ref{eq:disturbance})
by subtracting from the raw data
the normal gravity of the WGS84 reference ellipsoid ($\gamma(P)$)
using the formula of \citet{li2001a}.
Fig.~\ref{fig:sam-corrections}a show the calculated gravity disturbance of
South America.

We remove the gravitational effect of the topography
from the gravity disturbance
by modeling the ETOPO1 digital terrain model
\citep[][ \url{http://dx.doi.org/10.7289/V5C8276M}]{amante2009}
using tesseroids (Fig.~\ref{fig:sam-corrections}b).
We used the standard densities of $2670\ kg/m^3$ for continents and
$-1630\ kg/m^3$ for the oceans.
Fig.~\ref{fig:sam-corrections}c shows the calculated gravitational attraction
of the topographic masses at 50 km height.
Fig.~\ref{fig:sam-corrections}d shows the Bouguer disturbance
(Eq.~\ref{eq:bouguer}) obtained after subtracting the topographic effect from
the gravity disturbance.

The effect of sedimentary basins is removed using
tesseroid models of the three sedimentary layers present in the CRUST1.0 model
\citep[][ \url{http://igppweb.ucsd.edu/~gabi/rem.html}]{laske2013}.
Each sedimentary layer model includes the density
of each $1^\circ \times 1^\circ$ model cell.
Figs.~\ref{fig:sam-corrections}e-g show the thickness of the upper, middle, and
lower sedimentary layers, respectively.
The density-contrasts of the tesseroid model is obtained by subtracting
$2670\ kg/m^3$ from the density of each model element.
Fig.~\ref{fig:sam-corrections}h shows the combined gravitational attraction of
the sedimentary basin tesseroid model.
We subtract the total effect of sediments from the Bouguer disturbance in
Fig.~\ref{fig:sam-corrections}d to obtain
the sediment-free Bouguer disturbance (Fig.~\ref{fig:sam-data}a),
which will be used as input for the inversion.

Fig.~\ref{fig:sam-data}b shows the 937 known Moho depths (colored dots) which
were estimated from seismological data by \citet{assumpcao2013a}.
This data set is used in the validation procedure.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Inversion, cross-validation, and validation using seismological
data}

\begin{figure}
    \centering
    \includegraphics[width=0.5\columnwidth]{figures/south-america-cv}
    \caption{
        Cross-validation results for the South American Moho inversion.
        (a) Cross-validation to determine the regularization parameter $\mu$
        (Eq.~\ref{eq:goalfunction}).
        The minimum Mean Square Error (Eq.~\ref{eq:msemu}),
        shown as a red triangle,
        corresponds to $\mu = 10^{-10}$.
        (b) Validation to determine
        the reference level ($z_{ref}$) and the density-contrast ($\Delta\rho$).
        The colors represent the Mean Square Error (Eq.~\ref{eq:msehyper}).
        The minimum (red triangle) is found at $z_{ref} = 35\ km$
        and $\Delta\rho = 400\ kg/m^3$.
    }
    \label{fig:sam-cv}
\end{figure}

As in the CRUST1.0 synthetic data test (section~\ref{sec:crust1}),
we estimate the hyper-parameters in two steps.
First, we run the cross-validation to estimate
an optimal regularization parameter ($\mu$).
The starting estimate for all inversions is
60 km depth for all model parameters.
For this cross-validation,
we keep $z_{ref}$ and $\Delta\rho$ fixed to
$20\ km$ and $500\ kg/m^3$, respectively.
Second, we use the estimated $\mu$ to run the validation
using the seismological data of \citet{assumpcao2013a}
to estimate $z_{ref}$ and $\Delta\rho$,
thus obtaining the final estimated Moho depth model.

We split the sediment-free gravity disturbance (Fig.~\ref{fig:sam-data}a) into
the training and testing data sets.
The training data set is a regular grid with $0.4^\circ$ grid spacing
(twice the spacing of the original data grid)
and $N_{lat} \times N_{lon} = 201 \times 151$ grid points,
a total of 30,351 observations.
The remaining 90,350 points compose the testing data set.
We test 16 values of the regularization parameter ($\mu$)
equally spaced on a logarithmic scale between $10^{-10}$ and $10^{-2}$.
Fig.~\ref{fig:sam-cv}a shows the Mean Square Error (MSE)
as a function of $\mu$.
The minimum MSE is found at $\mu = 10^{-10}$, the lowest value of $\mu$ tested,
suggesting that little or no regularization is required.

We proceed with the validation using seismological data
using $\mu = 10^{-10}$ in all inversions.
We test all combinations of
nine values of $z_{ref}$, from 20 to 40 km with 2.5 km intervals,
and seven values of $\Delta\rho$, from 200 to 500 $kg/m^3$
with 50 $kg/m^3$ intervals.
Fig.~\ref{fig:sam-cv}b shows a pseudo-color map of the MSE
with respect to the \citet{assumpcao2013a} data set.
The MSE has a minimum, indicated by the red triangle,
at $z_{ref} = 35\ km$ and $\Delta\rho = 400\ kg/m^3$.
The minimum is not as well-defined as for the CRUST1.0 synthetic
(Fig.~\ref{fig:crust1-results}b),
which is expected because in reality $\Delta\rho$ is not homogeneous across all
of South America and the surrounding oceans.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Moho model for South America}

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figures/south-america-moho}
    \caption{
        The estimated Moho depth of South America.
        Dotted lines represent the boundaries between major geologic provinces
        \citep[after][]{assumpcao2013a, goutorbe2015};
        AD: Andean Province,
        AFB: Andean foreland basins,
        AM: Amazonas Basin,
        BR: Brazilian Shield,
        BO: Borborema province,
        CH: Chaco Basin,
        GB: Guyana Basin,
        GU: Guyana Shield,
        PB: Parnaíba Basin,
        PC: Parecis Basin,
        PR: Paraná Basin,
        PT: Patagonia province,
        SF: São Francisco Craton,
        SM: Solimões Basin.
        Solid orange lines mark the limits of the main lithospheric plates
        \citep{bird2003a};
        AF: Africa Plate,
        AN: Antarctica Plate,
        CA: Caribbean Plate,
        CO: Cocos Plate,
        SA: South America Plate,
        SC: Scotia Plate,
        NZ: Nazca Plate.
        The solid light grey line is the 35 km Moho depth contour.
    }
    \label{fig:sam-moho}
\end{figure*}

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figures/south-america-residuals}
    \caption{
        Residuals for the estimated South American Moho depth in
        Fig.~\ref{fig:sam-moho}.
        (a) Gravity residuals, defined as the difference between the observed data in Fig.~\ref{fig:sam-data}a
        and the data predicted by the estimate in Fig.~\ref{fig:sam-moho}.
        (b) Differences between the seismological depth estimates of
        \citet{assumpcao2013a} and our gravity-derived Moho depth estimate.
        The inset in b shows a histogram of the differences along with their
        calculated mean and standard deviation (std).
        Dotted lines mark the limits of major geologic provinces and
        lithospheric plates.
    }
    \label{fig:sam-residuals}
\end{figure*}

The final Moho depth model for South America is shown
as a pseudo-color map in Fig.~\ref{fig:sam-moho}.
The model is available in the online repository that accompanies
this contribution (see section~\ref{sec:software}).
Each model element is a $0.4^\circ \times 0.4^\circ$ tesseroid,
represented by the pixels in the pseudo-color map.

Our model differs significantly from CRUST1.0 (Fig.~\ref{fig:crust1-data}a)
but contains most of the large-scale features
present in the GMSA12 gravity-derived model of \citet{vandermeijde2013}.
The deepest Moho is along the central Andes, reaching depths upward of 70 km.
The oceanic areas present the shallowest Moho, ranging approximately
from 7.5 to 20 km.
The Brazilian and Guyana Shields have a deeper Moho (greater than 35 km),
with the deepest portions in the area around the São Francisco Craton
and the northern border of the Parecis Basin.
The Moho is shallower than 35 km along the Guyana Basin, the Andean foreland
basins, the Chaco Basin, and along the centers of the Solimões, Amazonas, and
Paraná Basins.

Fig.~\ref{fig:sam-residuals}a shows the gravity residuals,
defined as the difference between the observed and predicted gravity data.
Fig.~\ref{fig:sam-residuals}b shows the differences between
the seismic-derived Moho depths of \citet{assumpcao2013a}
(Fig.~\ref{fig:sam-data}b) and the depths of our gravity-derived model
(Fig.~\ref{fig:sam-moho}).
The differences shown in Fig.~\ref{fig:sam-residuals}b range from approximately
-23 to 23 km and have a mean of 1.18 km and a standard deviation of 6.84 km.
The gravity residuals and Moho depth differences from seismic are smallest in
the oceanic areas, southern Patagonia, and the eastern coast of the continent.
The largest gravity residuals are located along the Andes and correlate with
the deepest Moho depths.
These large residuals follow a pattern of a negative value in the center
flanked by positive values to the east and west.
This same pattern is observed in the CRUST1.0 synthetic test results
(Fig.~\ref{fig:crust1-results}).
In general, larger gravity residuals appear to be associated with sharp
variations in the estimated Moho depth.
Along the Andes, large differences with seismic data are correlated with the
larger gravity residuals.
Conversely, this correlation is absent from the large differences seen in the
Guyana, Paraná, and the Solimões Basins.
In the Borborema province, northeastern Brazil,
our model slightly overestimates the Moho depth.
On the other hand, our model underestimates the Moho depths in the Amazonas,
Solimões, and Paraná Basins.
Particularly in the Amazonas and Solimões Basins, where our model predicts a
Moho depth of approximately 30 km, the differences with the seismological
estimates can reach 10 km or more.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Discussion}

Differences between our Moho depth model and the seismological data
(Fig.~\ref{fig:sam-residuals}b)
may indicate regions where our initial assumptions
(summarized in Fig.~\ref{fig:anomalysketch})
are inadequate or where we have
failed to correct for all crustal and mantle sources.
The largest differences are seen along the Andean Province and
are likely caused by the fact that our model does not include
the subducting Nazca plate.
Furthermore, the CRUST1.0 synthetic data test (Fig.~\ref{fig:crust1-results})
suggests that our inversion method is not able to fully recover deep Moho
depths in the Andes,
even without the effect of the subducting plate.
In the Guyana Basin, our model is able to fit the gravity data
but differs from the seismological data by up to $\pm 10$ km with
no clear pattern for the distribution of the  differences.
A possible explanation is an inaccuracy in the CRUST1.0 sediment model
\citep{laske2013} used to correct our gravity data.
The inversion results will be biased
if the input data includes effects other than the anomalous Moho.

In the Amazon and Paraná Basins, our model fits the gravity data but
underestimates the seismological data by up to 15 km.
This indicates that a mass excess may be present in the crust or in the upper
mantle.
A body with positive density contrast whose gravitational effect was not
removed from the data during processing will make the observed gravity
disturbance greater than it would be otherwise.
This will cause the inversion to produce a shallower Moho estimate.
These discrepancies between gravity and seismological estimates have been noted
before by \citet{nunn1988} for the Amazon Basin and \citet{mariani2013}
for the Paraná Basin.
Both studies propose high density rocks in the lower crust as probable causes
for the discrepancies.
Another possible cause for the observed discrepancy in our model could be our
failure to fully remove from the data the effects of the igneous intrusions
present in both basins.
Using a sediment model for South America more detailed than CRUST1.0
might lead to more accurate results for these basins.

Greater confidence in our Moho model can be had in areas where it is able to
fit both the gravity and the seismological data.
In such places, a gravity-derived model can serve as an interpolator for
the seismological point estimates.
In general, our model fits both data sets in the oceans,
the Atlantic coast of the continent, and central Brazil.
In the Parnaíba Basin, our model has small differences with the few
seismological data points that fall inside and on the borders of the basin.
The Moho surrounding the basins continental borders is deep, following
inward with a shallower Moho, then a deeper step, and finally a shallower
part in the middle of the basin.
Recent deep seismic reflection studies by \citet{daly2014} that cross
the basin from west to east agree with our model.
The Northern border of the Parecis Basin has a deep Moho in our model that
is corroborated by a single seismological data point.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}

We have developed a computationally efficient gravity inversion method in
spherical coordinates.
Our method extends the Gauss-Newton formulation of Bott's method
to use tesseroids as model elements and Tikhonov
regularization.
The computational efficiency of our method is due to
using Bott's approximation for the Jacobian matrix
and using sparse matrix algorithms for arithmetic operations and the solution
of linear systems.
This approximation for the Jacobian matrix is adequate and the method converges
even when the data are at higher altitudes and the model is not outcropping,
as shown by the applications to synthetic data.
We employ hold-out cross-validation to estimate the regularization parameter
and a validation procedure using seismological data to estimate the Moho
density-contrast and the Normal Earth Moho depth.

There are two main advantages of the proposed method over previous works.
First, unlike the Parker-Oldenburg method or methods
using rectangular prisms, our inversion method does not require the data to be
projected onto a plane.
Second, the Parker-Oldenburg method and methods derived from Bott's method
cannot apply the traditional methodology for constraining inverse problems.
Our method has no such restriction because we use the formalism of traditional
Tikhonov inversion.
However, the proposed method is not without limitations.
It requires data on a regular grid and restricts the
model to be a regular mesh tied to the data grid.
Like Bott's method, our method only works for gravity disturbances and not for
the gravity gradients.

The test on simple synthetic data shows that our inversion method is able to
recover a smooth Moho relief with a homogeneous density-contrast distribution.
The inversion was not able to fully recover the shortest wavelength feature in
the model, possibly due to the smoothness constraints which tends to soften
high-frequency (sharp) variations.
The cross-validation Mean Square Error curve has a well-defined minimum,
indicating a value of the regularization parameter ($\mu$) whose corresponding
estimate best predicts data that were not included in the inversion.
Using this value of $\mu$ in the inversion leads to a stable solution
characterized by a smooth Moho relief with an acceptable data misfit.

The efficiency of the proposed method is because solving linear systems and
performing matrix multiplications together account for a mere 0.067\% of the
total computation time required for a single inversion.
The majority of the computation time (99.824\%) is spent on forward modeling.
Thus, we are able to retain the high computational efficiency of Bott's method
and use a classic Tikhonov regularization formulation.
This approach could, in theory, be extended to other types of regularization
(e.g., total variation) and misfit functions (e.g., re-weighted least squares)
already available in the literature.

The more complex synthetic data test based on CRUST1.0
shows that the validation using pointwise Moho depth information
is able to correctly estimate the density-contrast ($\Delta\rho$) and Normal
Earth Moho depth ($z_{ref}$).
This test indicates that the inversion neither correctly estimates Moho depth
nor adequately fits the gravity and pointwise data when sharp variations in
Moho depth occur.
This phenomenon is particularly strong in the region below the Andes.
A likely explanation is that the smoothness regularization
is intrinsically unable to produce sharp variations in Moho depth.
These effects might be mitigated with the use of sharpness-inducing
regularization, like
the weighted smoothness inversion,
Cauchy norm regularization,
entropic regularization,
total variation regularization,
or an adaptive mixed smoothness-sharpness regularization.

We applied the method proposed here to estimate the Moho depth for South
America.
Our estimated Moho depth model is in accordance with previous results.
The model fits well the gravity and seismic data in all oceanic regions, the
central portion of the Andean foreland, Patagonia, and coastal and central
parts of Brazil.
However, the model is unable to fit the gravity and seismic data in places with
sharp variations in Moho depth, particularly below the Andes and in the
boundaries of the main geotectonic provinces of the South American Plate, like
the Borborema province, the Parnaiba Basin, and the São Francisco Craton.
This might indicate that smoothness regularization should not be applied
indiscriminately to the whole model, as suggested by the CRUST1.0 synthetic
data test.
Another reason for the observed misfit might be the presence of crustal or
mantle density anomalies whose gravitational effects were not removed during
the data corrections.
In the Guyana Basin on the coastal region of Venezuela, along the central
Amazonas and Solimões Basins, and in the Paraná Basin, our Moho depth model is
able to fit the gravity data but differs significantly from the seismic data.
These discrepancies in the Paraná and Amazonas Basins are interpreted in the
literature as high density rocks in the lower crust.
In general, differences between a gravity and a seismically derived Moho model
may indicate the presence of crustal or mantle density anomalies that were
unaccounted for in the data processing.
Such locations warrant further detailed investigation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Acknowledgments}

We are indebted to the developers and maintainers of the open-source
software without which this work would not have been possible.
We thank Dr. Marcelo Assumpção for providing the seismological Moho depth
estimates,
Editor Dr. Gary Egbert and two anonymous reviewers for their thoughtful
comments that helped improve this work.
L. Uieda was supported by a scholarship from
Coordenação de Aperfeiçoamento de Pessoal de Nível Superior
(CAPES).
V.C.F. Barbosa was supported by a fellowship from
Conselho Nacional de Desenvolvimento Científico e Tecnológico (CNPq).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{gji}
\bibliography{biblio}

\end{document}